Application of a Cut-Cell Immersed Boundary Method for Wildfire Simulation over Complex Terrain

### 2020-06-09 Vanella_Article_Atmosphere_2020.tex

% Model equations, fluid side : scalar transport, momentum, energy. Stratification.

% Combustion, fuel, fire spread model equations.

@@

% How do we describe terrains - geometries. Methodology to get from DEM a geometry.

% What FDS does for Scalar Transport - 1 paragraph Modification close to the terrain surface: scalar transport, momentum, divergence.

% Connection with level set fire spread.

@@

@@

The process of steering the solution of the mass, momentum, and energy equations to match wind speed and direction gathered at random weather stations is known as \emph{data assimilation}~\cite{Kalnay:2003}. FDS uses a relatively simple data assimilation technique, a method known as \emph{nudging}, where you add a forcing term to the momentum equation to ``nudge'' the flow in the direction of the specified wind. FDS automatically drives the mean velocity components toward the desired values by adding a forcing term to the momentum equation. For example, the $u$ component equation is modified as follows:
\begin{equation}
   \frac{\partial u}{\partial t} + \ldots = \frac{u_0(z,t)-\overline{u}(z,t)}{\tau} \label{mean_forcing_u}
\end{equation}
Here, $u(\mathbf{x},t)$ is the computed velocity component, $u_0$ is the specified wind field component that can vary with height and time, and $\overline{u}$ is the average of the computed velocity component at the height $z$. The relaxation time scale, $\tau$, is an important parameter. The shorter this time scale, the faster the flow field will ``catch up'' to the mean wind, but at the expense of possibly washing out important flow structures. There is no firm guidance in the atmospheric modeling literature.

>>>>

For fire modeling on the gas phase, FDS employs a Low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}. Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\overline{p}(z,t)$ used on the equation of state (the ideal gas law is used), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.

Consider a mixture composed of $N$ chemical species $\alpha$, moving on a fixed point $\mathbf{x}$ in space with a mass weighted average velocity $\mathbf{u}(\mathbf{x},t)$. If $\rho(\mathbf{x},t)$ is the mixture density, the mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, were $\rho_\alpha(\mathbf{x},t)$ is the species density for the local thermodynamic conditions. The scalar transport and momentum equations take the form:
\begin{eqnarray}
   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \mathbf{J_{c}}_\alpha ) &=& - \nabla \cdot (\mathbf{J_{d}}_\alpha) + \dot{m}_\alpha'''  + 
    \dot{m}_{b,\alpha}'''\; , \; \alpha=1,\dots,N \label{eqn:spectran} \\
    \frac{\partial \mathbf{u}}{\partial t} - \mathbf{u} \times \boldsymbol{\omega} + \nabla H - \tilde{p} \, \nabla \left( 1/\rho\right) &=&
    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau} \right] \label{eqn:momtran}
\end{eqnarray}
where $\mathbf{J_{c}}_\alpha=\rho Y_\alpha  \mathbf{u}$, $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ are the convective and diffusive fluxes for component $\alpha$. Ficks Law for binary diffusion respect to a background species is assumed, $D_\alpha$ is the diffusivity of $\alpha$ respect to the background species, $\dot{m}_\alpha'''(\mathbf{x},t)$ is the combustion volumetric source for species $\alpha$ and $ \dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from sub-grid particles gasification. Also, in the momentum equation the head $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, and $\tilde{p}$ is the perturbation pressure, $\rho_0(z)$ is an equlibrium density, $\mathbf{g}=(0,0,g_z)$ gravity vector, $\mathbf{f}_{b}$ is the particle force term and $\boldsymbol{\tau}$ tensor accounting for molecular and sub-grid turbulent stresses parameterized by an eddy viscosity model. This last model is employed by default in large eddy simulation (LES) of outdoor fires. Details of the model equations and their derivation are found in reference~\cite{mcgratta_2013tr}.

Further, the flow field is subjected to the following divergence constraint:
%
\begin{eqnarray}
    ( \nabla \cdot \mathbf{u} )^{th} &=& 
    \left[ \frac{1}{\rho c_p T} - \frac{1}{\bar{p}} \right] 
    \frac{\partial \bar{p}}{\partial t} + \frac{w \rho_0 g_z}{\rho c_p T} \nonumber \\
    &+& \frac{1}{\rho c_p T} \left[ \dot{q}''' + \dot{q}_b''' - \nabla \cdot \dot{\mathbf{q}}'' - \mathbf{u} \cdot \nabla (\rho h_s) \right] \nonumber \\
    &+& \frac{1}{\rho} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right) \left[ \dot{m}_\alpha'''  + \dot{m}_{b,\alpha}''' - \nabla \cdot (\mathbf{J_{d}}_\alpha) - \mathbf{u} \cdot \nabla (\rho Y_\alpha) \right] \label{eq:divth}
\end{eqnarray}
%
where $h_s$ is the local sensible enthalpy, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $\dot{\mathbf{q}}''$ is the heat flux, sum of conduction, convection and radiation, and $T$ is the local temperature. The mixture specific heat at constant pressure and molecular weight are $c_p=\sum_{\alpha =1}^N{c_{p,\alpha} Y_\alpha}$ and $\overline{W}=\left(\sum_{\alpha =1}^N{Y_\alpha /W_\alpha} \right)^{-1}$ respectively, built from the individual species counterparts as defined. Finally, $w,\rho_0,g_z$ on the stratification term are respectively, the local vertical velocity, density for standard conditions (depends on vertical coordinate), and vertical gravity acceleration. This thermodynamic divergence is derived from and acts as a proxy for the sensible enthalpy evolution equation~\cite{mcdermo_2014}. Being $R$ the ideal gas constant, problem is closed with the ideal gas law $ \overline{p} = {\rho R T}/{\overline{W}}$.

% Fire spread model equations.
As commented in the introduction, the location of a fire front can be defined implicitly along the terrain surface through the zero of a two dimensional level set scalar function $\phi(x,y,t)$ in the form $\phi(x,y,t)=0$. This implicit form can be used to define and advection equation for the level set function:
%
\begin{equation}
   \frac{D \phi(x,y,t)}{D t} =  \frac{\partial \phi}{\partial t} + \mathbf{R}_{uv} \cdot \nabla (\phi) = 0 \label{eqn:lset}
\end{equation}
%
where $\mathbf{R}_{uv}=(R_u,R_v)$ is the rate of spread vector along the terrain plane. The magnitude of the rate of spread vector is computed by $|\mathbf{R}_{uv}|=R_0 \left(1 + \sqrt{(\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) \cdot (\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) } \right)$, being $R_0$ the zero velocity and zero slope rate of spread (baseline, dependent of fuel properties), and vectors $\boldsymbol{\phi}_W,\boldsymbol{\phi}_S \ge 0$ are computed using local wind, slope through empirical rules~\cite{Wilson:1980}. See Appendix 1 of reference~\cite{Bova:IJWF2015} for details on these calculations, as implemented in WFDS-LS and in this work.

@@

% How do we describe terrains - geometries. Methodology to get from DEM a geometry.
%
In figure~\ref{Fig:figure_1}a, a terrain is represented by its surface triangulation within an FDS cartesian mesh. Some cartesian cells and faces belonging to this mesh are transversed by the terrain surface. The remainder polyhedra and polygons that lay on the gas side of this intersected cartesian entities are called cut-cells and cut-faces. They discretize the fluid domain in the region surrounding the solid terrain. A detail of this unstructured mesh is. shown in figure~\ref{Fig:figure_1}b. The resulting cut-cells and cut-faces are of varied sizes and shapes, and situations where more than one cut-cell per cartesian cell are not uncommon. As described later, small cells pose a temporal stability constraint in explicit time integration schemes that needs to be addressed.  
It is noted that a two level of refinement grid hierarchy emerges. The coarse level is defined by the Cartesian entities, whereas the fine level is defined by the cut-cell or unstructured counterparts. Methods that solve discrete model equations on these grids are called cut-cell or embedded boundary methods~\cite{berger_2016}. Reliably defining the cut-cell mesh geometric properties and topology is a complex problem on itself, omitted here in order to keep the discussion succinct. 
%
\begin{figure}[h]
   \centering
   \includegraphics[trim = 12mm 42mm 8mm 25mm, clip,width=1\linewidth]{./figures/SketchFig1.pdf} 
   \put(-305,-5){(a)}    
   \put(-95,-4){(b)}
   \caption{Sketch of regular Cartesian and cut-cell grids around terrain: (a) An unstructured cut-cell region is defined on the gas phase side around a terrain $\mathbf{T}$. (b) Detail showing a slice representative of cut-cells and regular gas cells, terrain triangulation and two refinement level interpretation.}
   \label{Fig:figure_1}
\end{figure}  
%

\subsection{Scalar transport and energy discretization in unstructured region}

The equations for chemical species transport are discretized using the finite volume method (FV)~\cite{eymard_2000,leveque_2002}. See figure~\ref{Fig:figure_2}a. Integrating equation~\eqref{eqn:spectran} over a cut-cell $ii$ with control volume $\Omega_{ii}$:
%
\begin{equation}
 \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} d \Omega + \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)
      } d \Omega  = -\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \left(  \mathbf{J_{d}}_\alpha  \right)  } d \Omega + \int_{\Omega_{ii}} {( \dot{m}_\alpha'''+\dot{m}_{b,\alpha}''') } d \Omega \label{eq:intconvdiff}
\end{equation}
%
For a control volume that does not change with time, the time derivative and source terms are approximated by
%
\begin{eqnarray}
  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} d \Omega & = & \frac{\partial}{\partial t} \int_{\Omega_{ii}} {\rho Y_\alpha} d \Omega
  = \frac{\partial \widetilde{\rho \: Y_\alpha }_{ii}}{\partial t} V_{ii} \\
  \int_{\Omega_{ii}} {( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''')} d \Omega & = & (\widetilde{ \dot{m}_\alpha''' }_{ii}+ \widetilde{\dot{m}_{b,\alpha}'''}_{ii}) V_{ii} \label{eq:intcons}
\end{eqnarray}
%
where $V_{ii}$ is the volume of cell $ii$ and tildes imply cell averages. In the following the notation is simplified by dropping tildes, observing that quantities are cell or face averaged. Consider the FV discretization of the diffusive term (eqn.~\eqref{eq:intconvdiff}) on cut-cell $ii$ of figure~\ref{Fig:figure_2}a:
%
\begin{equation}
\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \left(  \mathbf{J_{d}}_\alpha  \right)  } d \Omega =
    \int_{\partial \Omega_{ii}} { \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right) \cdot \hat{\mathbf{n}}_{ii} } \: d \partial \Omega = \sum^{nf_c}_{k=1}
    \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \: A_k \label{eq:discfvdiffcc}
\end{equation}
%
The integral over the cut-cell volume has been transformed in an area integral on its $nf_c=5$ faces using the divergence theorem. As these $k$ faces with areas $A_k$ and normals out of the control volume $\hat{\mathbf{n}}_{ii,k}$ are planar by construction, the method for evaluation of their mean diffusive fluxes $\left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k$ will define the spatial accuracy of the discretization. A centroid to centroid (i.e. $\Delta x_{cc}$ in figure~\ref{Fig:figure_2}a) finite differences and linear interpolation are used to approximate $\boldsymbol{\nabla} Y_\alpha$ and $\rho D_\alpha$ for each face belonging to the gas phase. Also, a normal probe approach~\cite{balaras_2004} is employed to sample information from the fluid to define fluxes in boundary cut-faces. 
%
\begin{figure}[h]
   \centering
   \includegraphics[trim = 30mm 42mm 35mm 45mm, clip,width=1\linewidth]{./figures/SketchFig2.pdf} 
   \put(-335,-5){(a)}    
   \put(-95,-4){(b)}
   \caption{Sketch of cut-cell and faces: (a) Cut-cell $ii$ surrounded by gas phase regular and cut-faces (3-5), and boundary cut-faces (1-2).  (b) Interpolation sketch for wall modeled immersed boundary reconstruction of cut-face velocities.}
   \label{Fig:figure_2}
\end{figure}  
%

Similarly, the discretization of the advective term is:
%
\begin{equation}
    \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)} d \Omega =
     \int_{\partial \Omega_{ii}} { \left( \rho Y_\alpha \mathbf{u} \right) \cdot \hat{\mathbf{n}}_{ii} } \: d \partial \Omega =
     \sum^{nf_c}_{k=1} \left( \rho Y_\alpha \mathbf{u} \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \: A_k \label{eq:discfvadvcc}
\end{equation}
%
where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux limited interpolation to the cut-face~\cite{mcgratta_2013}. In the cut-cell region we use Godunov interpolation for the advective term. As the time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of equations~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known. The FV counterpart of the thermodynamic divergence expression for cut-cell $ii$ is:
%
%
\begin{eqnarray}
    ( \nabla \cdot \mathbf{u} )_{ii}^{th} \; V_{ii} &=&
    \left[ \frac{1}{(\rho c_p T)_{ii}} - \frac{1}{\bar{p}_{ii}} \right]
    \frac{\partial \bar{p}_{ii}}{\partial t} \; V_{ii} +
    \frac{w_{ii} \: \rho_0 g_z}{(\rho c_p T)_{ii}} V_{ii} \nonumber \\
    &+& \frac{1}{(\rho c_p T)_{ii}} \left[ (\dot{q}'''+ \dot{q}_b''')_{ii} V_{ii} -
    \sum_{k=1}^{nf_c} \dot{\mathbf{q}}''_{ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
    - \overline{\mathbf{u} \cdot \nabla (\rho h_s)} V_{ii} \right]  \\
    &+& \frac{1}{\rho_{ii}} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right)_{ii} \left[ (\dot{m}_\alpha'''+\dot{m}_{b,\alpha}''')_{ii} V_{ii} -
    \sum_{k=1}^{nf_c} \mathbf{J}_{\alpha,ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
    - \overline{\mathbf{u} \cdot \nabla (\rho Y_\alpha)} V_{ii} \right] \nonumber \label{eq:divth2}
\end{eqnarray}
%
where the over line terms refer to flux limited interpolation of corresponding scalars and terms defined with subscript $ii$ refer to cell defined quantities. All heat and mass fluxes and scalars are assumed known. Also, the vertical velocity $w_{ii}$ is interpolated to the cut-cell centroid. Details of the FDS time integration scheme can be found in references~\cite{mcgratta_2013,mcdermo_2014}. For each RK2 substep, the species transport equations are advanced in all FDS cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region. A similar procedure is done for the thermodynamic divergence. 
% Small cell linking for explicit time integration.
As an explicit time integrator is used, in general there will arise cut-cells whose small-size will penalize severely the time step. For these we link their unknown number to larger surrounding cells, which in turn defines mathematically an evolution equation for the union of both linked cells. Cell linking in the context of momentum equations can be found in reference~\cite{kirk_2003}.
In FDS, the momentum equations and pressure Poisson equation are solved in the cartesian mesh. Therefore, an immersed boundary method is used to reconstruct velocities in the cut-cell region. As will be shown next, a divergence integral equivalence argument is used to transfer the the divergence from cut-cells to under laying cartesian cells, and build the source of the Poisson equation for the pressure.

\subsection{Immersed boundary method on cut-cell region and wall modeling}

Collecting advective, shear stress and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within equation~\ref{eqn:momtran}, the model momentum transport problem can be casted as~\cite{mcgratta_2013}:
%
\begin{eqnarray}
  \frac{\partial \mathbf{u}(\mathbf{x},t)}{\partial t} &=& - \left[ \mathbf{F}(\mathbf{u},\mathbf{x},t) + \boldsymbol{\nabla} H(\mathbf{x},t) \right ]\label{eq:LowMachMom} \\
         \nabla \cdot \mathbf{u} (\mathbf{x},t) & = & \left(\nabla \cdot \mathbf{u} \right)^{th} \label{eq:LowMachDiv}
\end{eqnarray}
%
where equation~\eqref{eq:LowMachMom} is the momentum equation, subject to a specified divergence field, provided by equation~\eqref{eq:divth}. 
Boundary conditions are prescribed for $\mathbf{u}(\mathbf{x},t)$ on boundaries, including the immersed terrain.
%Classical fractional step methods for time integration of incompressible or low Mach flow are based on two operations: First, momentum transport to obtain intermediate velocities, and second, projection of velocities into a target divergence space. 
The previous equations are advanced in time using a fractional step method. As illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$. Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\nabla \cdot \mathbf{u}^{n+1} = \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th}$ known
%
\begin{eqnarray}
  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \left[ \mathbf{F}^n +  \boldsymbol{\nabla} H^n \right] \label{eq:LowMachMomEu}\\
  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th} \label{eq:LowMachDivEu}
\end{eqnarray}
%
where $\mathbf{u}^{n+1}$ represents a numerical solution at time $t_{n+1}$. This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator. As the potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is assumed responsible of enforcing the divergence condition and used on the projection step. Taking the divergence of equation~\eqref{eq:LowMachMomEu} and considering  the constraint~\eqref{eq:LowMachDivEu}, the two steps of the method are
%
\begin{enumerate}
  \item Solve Poisson equation for $H^n$:

\begin{equation}
   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
\end{equation}

  \item Obtain final velocity for step:

  \begin{equation}
     \mathbf{u}^{n+1} = \mathbf{u}^{n} - \Delta t \left[ \mathbf{F}^n +  \boldsymbol{\nabla} H^n \right] \label{it:FSProject}
   \end{equation}

   %The term $\hat{\mathbf{u}}^{n+1}=\mathbf{u}^{n} - \Delta t \mathbf{F}^n$ is known as intermediate velocity, and is a non divergence matching approximation to $\mathbf{u}^{n+1}$ (i.e. $\nabla \cdot \hat{\mathbf{u}}^{n+1} \neq \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th}$).
\end{enumerate}
%
A consequence of the projection scheme is that boundary conditions are required on the Poisson equation of step~\eqref{it:FSPoisson}. For explicit methods and stationary \textit{solid} boundaries, the corresponding boundary condition is \textit{homogeneous} Neumann for $H^n$ in $\partial \Omega,\partial \Omega_1,...,\partial \Omega_{nbods}$~\cite{perot_1993}.
Next, an approximation to the no-slip boundary condition at the immersed terrain surface is needed. To this end, a direct forcing immersed boundary method (IBM) for the momentum equations~\cite{fadlun_2000} is employed. A force field is computed on the discrete momentum equations on grid faces crossed by the immersed surfaces to approximate the no-slip boundary condition on these. In LES, the surrounding velocity field is modeled using an equilibrium boundary layer solution.

In figure~\ref{Fig:figure_2}b the velocity update in each of the gas phase cut-face centroids $d$ is done individualizing the point $B$ on the boundary and normal direction through these. Also, and external point $e_x$ through the normal $\mathbf{\hat{n}}$ is defined at a distance $\delta_{ex}$, of the order of the cartesian cells size. Known velocities and fluid parameters are interpolated from the surrounding fluid points $e_1,\dots,e_4$ to $e_x$. 
This information is used to estimate a target velocity at step $n+1$, $\mathbf{\hat{u}}_d^{n+1,k-1}$ at point $d$, assuming the log law equilibrium boundary layer solution~\cite{mcgratta_2013}. The target velocity $\hat{u}_d^{n+1,k-1}$ component on the cut-face centroid is flux matched to the under-laying cartesian face $E$ velocity component $\hat{u}_E^{n+1,k-1}$. Finally, an immersed boundary force 
%
\begin{equation}
F_E^k=-\frac{\hat{u}_E^{n+1,k-1}-u_E^{n}}{\Delta t} - \frac{\partial H^{n,k-1}}{\partial x}
\label{eq:FIBM}
\end{equation}
%
can be computed and used in equation~\eqref{it:FSPoisson} to take into account the presence of the body. The index $k$ refers in this context to the subiteration that can be performed in the IB force~\eqref{eq:FIBM} and projection~\eqref{it:FSPoisson}-\eqref{it:FSProject} to match final velocities with the wall modeled targets.

@@

% Coupling level set fire spread model with fluid solver.

@@

The process of steering the solution of the mass, momentum, and energy equations to match wind speed and direction gathered at random weather stations is known as \emph{data assimilation}~\cite{Kalnay:2003}. FDS uses a relatively simple data assimilation technique, a method known as \emph{nudging}, where you add a forcing term to the momentum equation to ``nudge'' the flow in the direction of the specified wind. FDS automatically drives the mean velocity components toward the desired values by adding a forcing term to the momentum equation. For example, the $u$ component equation is modified as follows:
\begin{equation}
   \frac{\partial u}{\partial t} + \ldots = \frac{u_0(z,t)-\overline{u}(z,t)}{\tau} \label{mean_forcing_u}
\end{equation}
Here, $u(\mathbf{x},t)$ is the computed velocity component, $u_0$ is the specified wind field component that can vary with height and time, and $\overline{u}$ is the average of the computed velocity component at the height $z$. The relaxation time scale, $\tau$, is an important parameter. The shorter this time scale, the faster the flow field will ``catch up'' to the mean wind, but at the expense of possibly washing out important flow structures.
